Interval-Valued Quantum Probabilities: Contextuality and the Born Rule

### 2017-05-20 prop_letter.tex
questions by GO
prop_letter.tex \hyphenation{cryp-to-de-ter-min-is-tic cryp-to-de-ter-min-ism} Added...

\usepackage{color}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{framed}
\newcommand{\amr}[1]{\begin{framed}\begin{minipage}{0.9\linewidth}\color{green}{Amr says: #1}\end{minipage}\end{framed}}
\newcommand{\yutsung}[1]{\begin{framed}\begin{minipage}{0.9\linewidth}\color{purple}{Yu-Tsung says: #1}\end{minipage}\end{framed}}
\newcommand{\set}[2]{\ensuremath{\left\{ {#1}~\middle|~{#2}\right\} }}
\newcommand{\Tr}{\ensuremath{\mathop{\mathrm{Tr}}\nolimits}}
\allowdisplaybreaks
\newcommand{\coreBorn}{\ensuremath{\mathrm{C}}}
\newcommand{\mul}[1][]{\ensuremath{\mu^{\mathrm{L{#1}}}}}
\newcommand{\mur}[1][]{\ensuremath{\mu^{\mathrm{R{#1}}}}}
\setcounter{secnumdepth}{3}
% https://tex.stackexchange.com/questions/345284/can-i-have-an-almost-non-breaking-space-in-latex
% Change non-breaking space to \nolinebreak[1] 
% Hence, it will not hyphenate too much surrounding words.
\newcommand{\nb}{\nolinebreak[1] }

@@

Our first observation is that essential objectives of our program to
achieve a computable theory of quantum mechanics and finite-resource
measurement, attempted previously with computable number
systems\nb\cite{usat,geometry2013apsrev4,DQT2014}, may be achievable by
extending classical interval-valued probability
measures\nb\cite{JamisonLodwick2004} to the quantum domain.  We begin by
axiomatizing a quantum interval-valued probability measure (QIVPM)
framework exploiting the fact that a density matrix approach can be
considered as equivalent to a description of quantum states using
probability measures\nb\cite{10.2307/2308516,Varadarajan2008}. %Mackey=10.2307
One result is that the expectation value of an observable with respect
to a QIVPM can be defined to be consistent with the constraints on the
expectation value of the same observable expected from continuous
quantum mechanics.  We then examine Gleason's
theorem\nb\cite{gleason1957,Redhead1987-REDINA,peres1995quantum} and the
Kochen-Specker theorem in the context of QIVPM's, and conclude that
they correspond to the limit of an infinite number of precise
intervals, and the set of crypto-deterministic intervals (the set\nb$\{\imposs,\necess\}$),
respectively.

@@

@@

Quantum IVPMs become classical IVPMs when the space of events includes
only mutually commuting sub-events, where $\events'\subseteq\events$
is called a sub-event space if it satisfies the following conditions.
\begin{itemize}
\item $\left\{ \mathbb{0},\mathbb{1}\right\} \subseteq\events'$.
\item For any projection\nb$P\in\events'$, we have $\mathbb{1}-P\in\events'$. 
\item For each pair of orthogonal projections\nb$P_0 \in \events'$ and
  $P_1 \in \events'$, we have $(P_0 + P_1) \in \events'$. 
\end{itemize}

@@

Our first result regarding QIVPM is that a QIVPM cannot be
crypto-deterministic in the sense of Peres\nb\cite{PeresRon1988,peres1995quantum}. We
will adapt the proof of the Kochen-Specker
theorem\nb\cite{kochenspecker1967,
  peres1995quantum,Redhead1987-REDINA,Griffiths2003} to show that it
is, in general, impossible to just have the two intervals \imposs\ and
\necess: every QIVPM must be parameterized by a collection of
intervals that includes at least one additional ``fuzzy'' interval and
some events must be associated with such uncertain intervals. The
significance of this observation is that the uncertainty in the
intervals is, in contrast to the classical case, \emph{a fundamental
  part of quantum theory}. There are in fact two sources of
uncertainty: one due to the finite measurement resources and one due
to complementarity which requires every quantum framework to be truly
probabilistic.

We will proceed by assuming a crypto-deterministic QIVMP $\bar{\mu}$
and derive a contradiction using the following nine ``Mermin-Peres square''
observables\nb\cite{Mermin1990Simple,peres1995quantum}:
\begin{center}
\begin{tabular}{|c|c|c|}
\hline 
$\mathbb{1}\otimes\sigma_{z}$  & $\sigma_{z}\otimes\mathbb{1}$  & $\sigma_{z}\otimes\sigma_{z}$ \tabularnewline
\hline 
$\sigma_{x}\otimes\mathbb{1}$  & $\mathbb{1}\otimes\sigma_{x}$  & $\sigma_{x}\otimes\sigma_{x}$ \tabularnewline
\hline 
$\sigma_{x}\otimes\sigma_{z}$  & $\sigma_{z}\otimes\sigma_{x}$  & $\sigma_{y}\otimes\sigma_{y}$ \tabularnewline
\hline 
\end{tabular}
\par\end{center}

@@

The expectation value is intuitively between two possible outcomes
which lie between the minimum and maximum bounds of the given
probability intervals, and this definition also consistent with the classical
Choquet integral on commuting sub-event space by adopting the classical
results\nb\cite{Rosenmuller1971,GilboaSchmeidler1994,Grabisch2016}.
If the QIVPM is crypto-deterministic, this
interval collapses to a singleton $\left\{ \lambda\right\} $, where
$\lambda$ is an eigenvalue of $\mathbf{O}$. If we define the
multiplication of two singleton sets\nb$\left\{ \lambda_{1}\right\} $
and $\left\{ \lambda_{2}\right\} $ to be
$\left\{ \lambda_{1}\lambda_{2}\right\} $, then we have
\begin{equation}
\prod_{j=1}^{N}\expval{\mathbf{O}_{j}}_{\bar{\mu}}=\expval{\prod_{j=1}^{N}\mathbf{O}_{j}}_{\bar{\mu}}\label{eq:Kochen-Specker-Rules}
\end{equation}
for any sequence of commuting observables\nb$\left\{ \mathbf{O}_{j}\right\} _{j=1}^{N}$.

@@

To adapt the axioms of such IVPMs to the quantum case, we proceed as
follows. First we fix a collection of intervals $\mathscr{I}$ that
represents an abstraction of the available experimental resources. At
one extreme, infinitely precise resources would correspond to the
collection of intervals $\{[x,x] \mid x \in [0,1]\}$. This case should
coincide with the idealized situation above. At another extreme, an
experiment with ``minimal'' resources would correspond to the
following three intervals $\imposs=\left[0,0\right]$,
$\necess=\left[1,1\right]$, and $\unknown=[0,1]$. As before, the first
two intervals are associated with the empty projection and the
identity projection but every other projection would be associated
with $\unknown$; its probability can be anything in the range
$[0,1]$. A third special situation would be to consider IVPMs with
only the two intervals $\imposs=\left[0,0\right]$ and
$\necess=\left[1,1\right]$. This is what Peres calls the
\emph{crypto-deterministic} case in which there is no ambiguity about
the status of any event; we know for certain if it will occur or not.

>>>>

\usepackage{color}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{framed}
\newcommand{\amr}[1]{\begin{framed}\begin{minipage}{0.9\linewidth}\color{green}{Amr says: #1}\end{minipage}\end{framed}}
\newcommand{\yutsung}[1]{\begin{framed}\begin{minipage}{0.9\linewidth}\color{purple}{Yu-Tsung says: #1}\end{minipage}\end{framed}}
\newcommand{\set}[2]{\ensuremath{\left\{ {#1}~\middle|~{#2}\right\} }}
\newcommand{\Tr}{\ensuremath{\mathop{\mathrm{Tr}}\nolimits}}
\allowdisplaybreaks
\newcommand{\coreBorn}{\ensuremath{\mathrm{C}}}
\newcommand{\mul}[1][]{\ensuremath{\mu^{\mathrm{L{#1}}}}}
\newcommand{\mur}[1][]{\ensuremath{\mu^{\mathrm{R{#1}}}}}
\setcounter{secnumdepth}{3}
% https://tex.stackexchange.com/questions/345284/can-i-have-an-almost-non-breaking-space-in-latex
% Change non-breaking space to \nolinebreak[1] 
% Hence, it will not hyphenate too much surrounding words.
\newcommand{\nb}{\nolinebreak[1] }
\hyphenation{cryp-to-de-ter-min-is-tic cryp-to-de-ter-min-ism}

@@

Our first observation is that essential objectives of our program to
achieve a computable theory of quantum mechanics and finite-resource
measurement, attempted previously with computable number
systems\nb\cite{usat,geometry2013apsrev4,DQT2014}, may be achievable by
extending classical interval-valued probability
measures\nb\cite{JamisonLodwick2004} to the quantum domain.  We begin by
axiomatizing a quantum interval-valued probability measure (QIVPM)
framework exploiting the fact that a density matrix approach can be
considered as equivalent to a description of quantum states using
probability measures\nb\cite{10.2307/2308516,Varadarajan2008}. %Mackey=10.2307
One result is that the expectation value of an observable with respect
to a QIVPM can be defined to be consistent with the constraints on the
expectation value of the same observable expected from continuous
quantum mechanics.  We then examine Gleason's
theorem\nb\cite{gleason1957,Redhead1987-REDINA,peres1995quantum} and the
Kochen-Specker theorem in the context of QIVPM's, and conclude that
they correspond to the limit of an infinite number of precise
intervals, and the set of cryptodeterministic intervals (the set\nb$\{\imposs,\necess\}$),
respectively.

@@

A question we want to entertain in this paper is \ldots
Can one incorporate in a rigorous mathematical framework the idea of
finite precision measurement? Yes QIVPM. Now what is the fate of the 
foundational theorems of QM when this framework is applied?

@@

Quantum IVPMs become classical IVPMs when the space of events includes
only mutually commuting events, where $\events'\subseteq\events$
is called a sub-event space if it satisfies the following conditions.
\begin{itemize}
\item $\left\{ \mathbb{0},\mathbb{1}\right\} \subseteq\events'$.
\item For any projection\nb$P\in\events'$, we have $\mathbb{1}-P\in\events'$. 
\item For each pair of orthogonal projections\nb$P_0 \in \events'$ and
  $P_1 \in \events'$, we have $(P_0 + P_1) \in \events'$. 
\end{itemize}

@@

Our first result regarding QIVPM is that a QIVPM cannot be
cryptodeterministic in the sense of Peres\nb\cite{PeresRon1988,peres1995quantum}. We
will adapt the proof of the Kochen-Specker
theorem\nb\cite{kochenspecker1967,
  peres1995quantum,Redhead1987-REDINA,Griffiths2003} to show that it
is, in general, impossible to just have the two intervals \imposs\ and
\necess: every QIVPM must be parameterized by a collection of
intervals that includes at least one additional ``fuzzy'' interval and
some events must be associated with such uncertain intervals. The
significance of this observation is that the uncertainty in the
intervals is, in contrast to the classical case, \emph{a fundamental
  part of quantum theory}. There are in fact two sources of
uncertainty: one due to the finite measurement resources and one due
to complementarity which requires every quantum framework to be truly
probabilistic.

We will proceed by assuming a cryptodeterministic QIVMP $\bar{\mu}$
and derive a contradiction using the following nine ``Mermin-Peres square''
observables\nb\cite{Mermin1990Simple,peres1995quantum}:
\begin{center}
\begin{tabular}{|c|c|c|}
\hline 
$\mathbb{1}\otimes\sigma_{z}$  & $\sigma_{z}\otimes\mathbb{1}$  & $\sigma_{z}\otimes\sigma_{z}$ \tabularnewline
\hline 
$\sigma_{x}\otimes\mathbb{1}$  & $\mathbb{1}\otimes\sigma_{x}$  & $\sigma_{x}\otimes\sigma_{x}$ \tabularnewline
\hline 
$\sigma_{x}\otimes\sigma_{z}$  & $\sigma_{z}\otimes\sigma_{x}$  & $\sigma_{y}\otimes\sigma_{y}$ \tabularnewline
\hline 
\end{tabular}
\par\end{center}

@@

The expectation value is intuitively between two possible outcomes
which lie between the minimum and maximum bounds of the given
probability intervals, and this definition also consistent with the classical
Choquet integral on commuting sub-event space by adopting the classical
results\nb\cite{Rosenmuller1971,GilboaSchmeidler1994,Grabisch2016}.
If the QIVPM is cryptodeterministic, this
interval collapses to a singleton $\left\{ \lambda\right\} $, where
$\lambda$ is an eigenvalue of $\mathbf{O}$. If we define the
multiplication of two singleton sets\nb$\left\{ \lambda_{1}\right\} $
and $\left\{ \lambda_{2}\right\} $ to be
$\left\{ \lambda_{1}\lambda_{2}\right\} $, then we have
\begin{equation}
\prod_{j=1}^{N}\expval{\mathbf{O}_{j}}_{\bar{\mu}}=\expval{\prod_{j=1}^{N}\mathbf{O}_{j}}_{\bar{\mu}}\label{eq:Kochen-Specker-Rules}
\end{equation}
for any sequence of commuting observables\nb$\left\{ \mathbf{O}_{j}\right\} _{j=1}^{N}$.

@@

To adapt the axioms of such IVPMs to the quantum case, we proceed as
follows. First we fix a collection of intervals $\mathscr{I}$ that
represents an abstraction of the available experimental resources. At
one extreme, infinitely precise resources would correspond to the
collection of intervals $\{[x,x] \mid x \in [0,1]\}$. This case should
coincide with the idealized situation above. At another extreme, an
experiment with ``minimal'' resources would correspond to the
following three intervals $\imposs=\left[0,0\right]$,
$\necess=\left[1,1\right]$, and $\unknown=[0,1]$. As before, the first
two intervals are associated with the empty projection and the
identity projection but every other projection would be associated
with $\unknown$; its probability can be anything in the range
$[0,1]$. A third special situation would be to consider IVPMs with
only the two intervals $\imposs=\left[0,0\right]$ and
$\necess=\left[1,1\right]$. This is what Peres calls the
\emph{cryptodeterministic} case in which there is no ambiguity about
the status of any event; we know for certain if it will occur or not.
